---
title: "KFF report"
author: "Stas Kolenikov"
date: "`r Sys.Date()`"
params:
  month: April
  year:  2019
output:
  html_document:
    df_print: paged
    code_folding: show
  word_document:
    reference_docx: Abt_2018_Report_Template.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r libraries, message=FALSE, warning=FALSE}
# to read SPSS files
library(haven)
# to process data
library(tidyverse)
# to produce neat tables
library(knitr)
library(kableExtra)
library(flextable)
# to abstract directories
library(here)
```

```{r deff_function}
DEFF <- function(x) ( sum(x*x)*nrow(x)/(sum(x)*sum(x)) )
# checks
stopifnot( DEFF(c(1,1,1)) == 1)
stopifnot( DEFF(c(10,10,10)) == 1)
stopifnot( DEFF(c(5,0,0)) == 5)
neff <- function(x) nrow(x)/DEFF(x)
# checks
stopifnot( neff(c(1,1,1)) == 3)
stopifnot( neff(c(10,10,10)) == 3)
stopifnot( neff(c(5,0,0)) == 1)
```

```{r read_spss}
KFF_data <- haven::read_spss(here("data",paste0("Kaiser Health Tracker Poll_",
                                                params$month,params$year,"_Final.sav")))
```

## Sample report

```{r svydates}
# need to preprocess the survey dates from the intvwdate variable
KFF_data %>% mutate(idate = as.Date(intvwdate,format="%y%m%d")) -> KFF_data
idate_start <- min(KFF_data$idate)
idate_end <- max(KFF_data$idate)
```

This KFF Health Tracking Poll was designed and analyzed 
by public opinion researchers at the Kaiser Family Foundation (KFF). 
The survey was conducted from `r format(idate_start,"%B %d, %Y")` to `r format(idate_end,"%B %d, %Y")`, 
among a nationally representative random digit dial telephone sample of 
`r format(nrow(KFF_data), big.mark=",")` adults ages 18 and older, living in the United States, 
including Alaska and Hawaii (note: persons without a telephone could not be included 
in the random selection process). The sample included 
`r KFF_data %>% filter(sample07 %in% c(10,11)) %>% summarize(n=n()) %>% unlist()` respondents 
reached by calling back respondents that had previously completed an interview 
on the KFF Tracking poll more than nine months ago. 
Computer-assisted telephone interviews conducted 
by landline (`r KFF_data %>% filter(sample==1) %>% summarize(n=n()) %>% unlist()`) 
and cell phone (`r KFF_data %>% filter(sample==2) %>% summarize(n=n()) %>% unlist()`, 
including `r KFF_data %>% filter(iphoneuse==3) %>% summarize(n=n()) %>% unlist()` 
who had no landline telephone) were carried out in English 
(`r KFF_data %>% filter(lang==1) %>% summarize(n=n()) %>% unlist()`)
and Spanish (`r KFF_data %>% filter(lang==2) %>% summarize(n=n()) %>% unlist()`).

### MOE table

```{r moe_data}
KFF_MOE <- rbind(
  # total
  tibble(Group="Total", unweighted_n = nrow(KFF_data), 
         neff = (KFF_data %>% filter(TRUE) %>% select(weight_ssrs) %>% as.vector() %>% neff() )
         ),
  # blank line
  tibble(Group="Party ID", unweighted_n = NA, neff = NA),
  # party groups
  tibble(Group="Democrats", unweighted_n = nrow(KFF_data %>% filter(party5 == 1)), 
         neff = (KFF_data %>% filter(party5 == 1) %>% select(weight_ssrs) %>% as.vector() %>% neff() )
         ),
  tibble(Group="Republicans", unweighted_n = nrow(KFF_data %>% filter(party5 == 5)), 
         neff = (KFF_data %>% filter(party5 == 5) %>% select(weight_ssrs) %>% as.vector() %>% neff() )
         ),
  tibble(Group="Independents", unweighted_n = nrow(KFF_data %>% filter(party5 %in% c(2,3,4))), 
         neff = (KFF_data %>% filter(party5 %in% c(2,3,4)) %>% select(weight_ssrs) %>% as.vector() %>% neff() )
         ),
  tibble(Group="Dem-leaning Indeps", unweighted_n = nrow(KFF_data %>% filter(party5 == 2)), 
         neff = (KFF_data %>% filter(party5 == 2) %>% select(weight_ssrs) %>% as.vector() %>% neff() )
         ),
  tibble(Group="Rep-leaning Indeps", unweighted_n = nrow(KFF_data %>% filter(party5 == 4)), 
         neff = (KFF_data %>% filter(party5 == 4) %>% select(weight_ssrs) %>% as.vector() %>% neff() )
         )
)
KFF_MOE %>% mutate(
  `N (unweighted)` = if_else(is.na(unweighted_n), "", format(unweighted_n, big.mark=",") ),
  MOE_num = qnorm(0.975)*sqrt(0.5*0.5/neff),
  MOE = if_else( is.na(MOE_num), "", paste0("+/- ", round(100*MOE_num,1), " percentage points") )
  ) -> KFF_MOE
saveRDS(KFF_MOE,file=here('KFF_MOE.Rds'))
```

Numbers of respondents and margins of sampling error for key subgroups are shown in the table below. 
For results based on other subgroups, the margin of sampling error may be higher. 

```{r MOE_table}
KFF_MOE %>% select(Group, `N (unweighted)`,MOE) -> KFF_MOE_table
if (is_html_output()) { 
  KFF_MOE_table %>%  kable() %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
} else KFF_MOE_table
```

